|

### 2023-01-29 paper.Rmd
Sun Jan 29 12:00:04 CET 2023: update-remote.sh

<!-- LaTeX Setup -->
\setcounter{page}{1}
\tableofcontents
\thispagestyle{empty}
\newpage
\listoftables
\listoffigures
\thispagestyle{empty}
\newpage

@@

Besides studying the performance for the airfoil dataset, additional real world symbolic regressions experiments used the datasets[@machine_learning_repo] described in table \@ref(tab:full-run-realWorldSymReg-problems).

# Experiments

@@

Figure \@ref(fig:first-gen-airfoil-byNeurons) summarizes the median reconstruction error for a variable number of hidden neurons. As expected, increasing the number of hidden neurons leads to stronger overfitting to the training dataset which results in small training errors but growing testing errors. For the experiments run with 50 and 100 hidden neurons, a small benefit of using pre-training can be observed. Using only 50 hidden neurons leads to a smaller gap between training and testing error for the pre-trained DAE-LSTM. For 100 hidden neurons, regular DAE-GP not only starts to overfit to the training data earlier than pre-trained DAE-GP, also it can be seen that pre-trained DAE-GP starts to overfit at a lower reconstruction error which would result in a higher model performance if training strategies such as early stopping would be employed.

@@

@@

The first interesting observation from this experiment is, that the pre-trained DAE-GP shows slightly lower mean and median performance than regular DAE-GP for both the training and the testing set in overall fitness. Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behaviour.

```{r airfoil-median-fitness-30gens, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Airfoil (2HL)", out.width="90%"}
knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens_withGP/mean_median_fitness_byGens.png")
```
The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as boxplots in figure  \@ref(fig:airfoil-distribution-fitness-30gens). The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP on both sets since it shows a lower median RMSE as well as less dispersion of the solutions measured by the standard deviation.

@@

<!-- Overview Real World SymReg  -->

To gather more confidence in the previous results, the pre-training strategy for DAE-GP was tested for three additional real world symbolic regression problems taken from @machine_learning_repo.^[see table \@ref(tab:full-run-realWorldSymReg-problems)]

>>>>

<!-- LaTeX Setup -->
\setcounter{page}{1}
<!-- \tableofcontents -->
\thispagestyle{empty}
\newpage
\listoftables
\listoffigures
\thispagestyle{empty}
\newpage

@@

# Results for the Airfoil Dataset

@@

Figure \@ref(fig:first-gen-airfoil-byNeurons) summarizes the median reconstruction error for a variable number of hidden neurons. As expected, increasing the number of hidden neurons leads to stronger overfitting to the training dataset which results in small training errors but growing testing errors. For the experiments run with 50 and 100 hidden neurons, a small benefit of using pre-training can be observed. Using only 50 hidden neurons leads to a smaller gap between training and testing error for the pre-trained DAE-LSTM. For 100 hidden neurons, regular DAE-GP not only starts to overfit to the training data earlier than pre-trained DAE-GP, also it can be seen that pre-trained DAE-GP starts to overfit at a lower reconstruction error which would result in a higher model performance if training strategies such as early stopping would be employed. Doubling the number of hidden neurons to 200 results in very strong overfitting behaviour for both DAE-GP variants starting around the 200.th epoch.

@@

The reults for the experiment with varying number of hidden layers are summarized in figure \@ref(fig:first-gen-airfoil-byLayers). Again, it can be observed that by increasing the dimension of the DAE-LSTM network both algorithms increasingly overfit to the training data. While for a single hidden layer overfitting begins at around the 400th. epoch, doubling the number of hidden layers to two leads to both algorithms overfitting at around the 200.th epoch. In general over all three sub experiments, both algorithms start to overfit after approximately the same number of epochs but pre-training leads to reaching the point of overfitting at a higher reconstruction error which would result in a inferrior model performance if the early stopping training strategy would be used during model training.

@@

The first interesting observation from this experiment is, that the pre-trained DAE-GP shows slightly lower mean RMSE (higher fitness) but a higher median RMSE (lower fitness) than regular DAE-GP for both the training and the testing set in overall fitness.Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behavior.

```{r airfoil-median-fitness-30gens, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Airfoil (2HL)", out.width="90%"}
knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens_withGP/mean_median_fitness_byGens.png")
```
The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as boxplots in figure  \@ref(fig:airfoil-distribution-fitness-30gens). The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP on both sets since it shows a lower median RMSE as well as less dispersion of the solutions measured by the standard deviation. An interesting observation from studying the distribution of final fitness scores for both algorithms is that for pre-trained DAE-GP the results

@@

To gather more confidence in the previous results, the pre-training strategy for DAE-GP was tested for three additional real world symbolic regression problems taken from @machine_learning_repo. Table \@ref(tab:full-run-realWorldSymReg-problems) briefly summarizes the benchmark problems.

### 2023-01-29 paper.tex
Sun Jan 29 12:00:04 CET 2023: update-remote.sh

{
\setcounter{tocdepth}{3}
\tableofcontents
}
\setcounter{page}{1}
\tableofcontents
\thispagestyle{empty}
\newpage
\listoftables
\listoffigures
\thispagestyle{empty}
\newpage

@@

Besides studying the performance for the airfoil dataset, additional real world symbolic regressions experiments used the datasets{[}\protect\hyperlink{ref-machine_learning_repo}{2}{]} described in table \ref{tab:full-run-realWorldSymReg-problems}.

\hypertarget{results-airfoil-dataset}{%
\section{Results Airfoil Dataset}\label{results-airfoil-dataset}}

@@

For each configuration I tested traditional DAE-GP as well as a pre-trained DAE-GP resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=\(120\)). To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{13}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =20000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\). The training of DAE-LSTM uses a fixed number of 1000 epochs until termination and uses Adam optimization for gradient descent. The reason for using a high amount of 1000 fixed training epochs is to deliberately force the DAE-LSTM to overfit to the training data.

@@

@@

\caption{First Generation Median Training Error for variable number of hidden Neurons}\label{fig:first-gen-airfoil-byNeurons}
\end{figure}

@@

\caption{First Generation Median Training Error for variable number of hidden Layers}\label{fig:first-gen-airfoil-byLayers}
\end{figure}

@@

To gather more confidence in the previous results, the pre-training strategy for DAE-GP was tested for three additional real world symbolic regression problems taken from {[}\protect\hyperlink{ref-machine_learning_repo}{2}{]}.\footnote{see table \ref{tab:full-run-realWorldSymReg-problems}}

>>>>

{
\setcounter{tocdepth}{3}
\tableofcontents
}
\setcounter{page}{1}

\thispagestyle{empty}
\newpage
\listoftables
\listoffigures
\thispagestyle{empty}
\newpage

@@

\hypertarget{results-for-the-airfoil-dataset}{%
\section{Results for the Airfoil Dataset}\label{results-for-the-airfoil-dataset}}

@@

For each configuration I tested traditional DAE-GP as well as a pre-trained DAE-GP resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=\(120\)). To avoid creating biased results through the presence of sampling error inside the pre-training population (see {[}\protect\hyperlink{ref-sampling_err_gp}{13}{]}), the population size for the pre-training phase is chosen very high with the size of \(\hat{P}\) =10000 where 50\% of \(\hat{P}\) is used for the training population \(\hat{P}_{train}\) and the remaining 50\% are used for the hold-out validation population \(\hat{P}_{test}\). The training of DAE-LSTM uses a fixed number of 1000 epochs until termination and uses Adam optimization for gradient descent. The reason for using a high amount of 1000 fixed training epochs is to deliberately force the DAE-LSTM to overfit to the training data.

@@

Figure \ref{fig:first-gen-airfoil-byNeurons} summarizes the median reconstruction error for a variable number of hidden neurons. As expected, increasing the number of hidden neurons leads to stronger overfitting to the training dataset which results in small training errors but growing testing errors. For the experiments run with 50 and 100 hidden neurons, a small benefit of using pre-training can be observed. Using only 50 hidden neurons leads to a smaller gap between training and testing error for the pre-trained DAE-LSTM. For 100 hidden neurons, regular DAE-GP not only starts to overfit to the training data earlier than pre-trained DAE-GP, also it can be seen that pre-trained DAE-GP starts to overfit at a lower reconstruction error which would result in a higher model performance if training strategies such as early stopping would be employed. Doubling the number of hidden neurons to 200 results in very strong overfitting behaviour for both DAE-GP variants starting around the 200.th epoch.

@@

\caption{First Generation Median Reconstruction Error for variable number of hidden Neurons}\label{fig:first-gen-airfoil-byNeurons}
\end{figure}

The reults for the experiment with varying number of hidden layers are summarized in figure \ref{fig:first-gen-airfoil-byLayers}. Again, it can be observed that by increasing the dimension of the DAE-LSTM network both algorithms increasingly overfit to the training data. While for a single hidden layer overfitting begins at around the 400th. epoch, doubling the number of hidden layers to two leads to both algorithms overfitting at around the 200.th epoch. In general over all three sub experiments, both algorithms start to overfit after approximately the same number of epochs but pre-training leads to reaching the point of overfitting at a higher reconstruction error which would result in a inferrior model performance if the early stopping training strategy would be used during model training.

@@

\caption{First Generation Median Reconstruction Error for variable number of hidden Layers}\label{fig:first-gen-airfoil-byLayers}
\end{figure}

@@

To gather more confidence in the previous results, the pre-training strategy for DAE-GP was tested for three additional real world symbolic regression problems taken from {[}\protect\hyperlink{ref-machine_learning_repo}{2}{]}. Table \ref{tab:full-run-realWorldSymReg-problems} briefly summarizes the benchmark problems.
